```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
library(glue)
```

# Open Burning - Yard Waste {#yardwaste}

## Sector Descriptions and Overview 

This source category covers intentional burning for waste disposal purposes of leaf and brush species in outdoor areas. Emission estimates for leaf and brush waste burning are a function of the amount of waste burned per year. Table \@ref(tab:yardwaste-sccs) shows the nonpoint SCCs in the NEI for open burning of yard waste, whether generated by EPA, or provided by SLTs.

```{r, include=FALSE}
df <- read_excel("./tables/Section-36/yardwaste-sccs.xlsx")
df[is.na(df)] <- ""
```

```{r yardwaste-sccs, tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Source Classification Codes used in the for open burning of yard waste.',
  format.args = list(scientific = FALSE, trim = TRUE)
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("hold_position", "scale_down"))
```

## EPA-developed estimates

The calculations for estimating the emissions from the burning of yard waste first involve estimating the amount of leaf and brush waste generated in each county. The amount of county-level yard waste burned is estimated by calculating the per capita amount of leaf and brush waste generated and multiplying that by the number of people likely to burn waste in each county. The number of people likely to burn waste is based on the rural population in each county. The total amount of yard waste burned is multiplied by emissions factors for criteria air pollutants (CAPs) and hazardous air pollutants (HAPs) to estimate emissions of these pollutants from yard waste burning. 

### Activity Data

The activity data for this source category is the amount of leaf and brush waste generated, which is estimated using data from the EPA report Advancing Sustainable Materials Management [[ref 1](#yardwaste-references)]. The report presents the total mass of waste generated from the residential and commercial sectors in the United States by type of waste for the calendar year 2018.

The per capita value of yard waste subject to burning was developed based on EPA’s total amount of waste generated [[Table 1 in ref 1](#yardwaste-references)]. According to the 2010 version of the same EPA report, residential waste generation accounts for 55-65% of the total waste from the residential and commercial sectors [[ref 2](#yardwaste-references)]; for the per capita calculation, the median value of 60% of total waste generated is assumed. This number is multiplied by the amount of yard waste generated and divided by the U.S. population in 2018 [[ref 3](#yardwaste-references)] to determine the per capita amount of yard waste generated in the United States. 

\begin{equation} 
  PC_{yw} = \frac{YW \times 0.60}{P_{US}}
  (\#eq:yardwaste-per-capita)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( PC_{yw} \) = Per-capita value of yard waste generated in the US, in tons per person per year <br> \newline
  \( YW \) = Annual yard waste generated, in million tons <br> \newline
  \( P_{US} \) = US population  <br> \newline
</div>

Open burning of yard waste is generally not practiced in urban areas. Therefore, only the rural population in each county is assumed to practice open burning. The rural and urban populations are taken from U.S. Census data for each county [[ref 3](#yardwaste-references)]. It is assumed that 24% of the rural population burns yard waste [[ref 4](#yardwaste-references)].

\begin{equation}
  PBurn_{c} = RPop_{c} \times 0.24
  (\#eq:yardwaste-rural-frac)
\end{equation}

<div style="margin-left: 2em;">
  Where: <br> \newline
  \(PBurn_{c} \) = Population likely to burn yard waste in county *c* <br>\newline
  \(RPop_{c} \) = Rural population in county *c* <br>\newline
</div>

The number of people likely to burn waste in each county (from Equation \@ref(eq:yardwaste-rural-frac)) is then used with the values of per-capita yard waste generated to determine the amount of household yard waste burned. The first assumption concerns the composition of yard waste; of the total amount of yard waste generated, yard waste composition is assumed to be 25 percent leaves, 25 percent brush, and 50 percent grass by weight [[ref 5](#yardwaste-references)]. However, open burning of grass clippings is not typically practiced by homeowners, and as such only estimates for leaf burning and brush burning are developed. 

The second assumption adjusts for variations in vegetation; the percentage of forested acres (including rural forest and urban forest) is determined using Version 2 of the Biogenic Emission Landuse Database (BELD2) within the Biogenic Emissions Inventory System (BEIS) [[ref 6](#yardwaste-references)]. Based on this percentage, county-level yard waste values are adjusted according to the values in Table \@ref(tab:yardwaste-adjfactors). To better account for the native vegetation that likely occurs in residential yards of farming states, agricultural land acreage is subtracted before calculating the percentage of forested acres. All municipios in Puerto Rico and counties in the U.S. Virgin Islands, Hawaii, and Alaska were assumed to have greater than 50 percent forested acres.

\begin{equation}
  LW_{c} = PBurn_{c} \times PC_{yw} \times YWFr_{t} \times AF_{fa,c}
  (\#eq:leaf-waste-burned)
\end{equation}

\begin{equation}
  BW_{c} = PBurn_{c} \times PC_{yw} \times YWFr_{t} \times AF_{fa,c}
  (\#eq:brush-waste-burned)
\end{equation}

<div style="margin-left: 2em;">
  Where: <br> \newline
  \(LW_{c} \) = Annual leaf waste burned in county *c*, in tons <br>\newline
  \(BW_{c} \) = Annual brush waste burned in county *c*, in tons <br>\newline
  \(PBurn_{c} \) = Population likely to burn yard waste in county *c* <br>\newline
  \(PC_{yw} \) = Per-capita value of yard waste in the U.S., in tons per person <br>\newline
  \(YWFr_{t} \) = Fraction of total yard wate for waste type *t* (leaf or brush) <br>\newline
  \(AF_{fa,c} \) = Adjustment factor based on percent of forested acres in county *c* <br>\newline
</div>

```{r, include=FALSE}
df <- read_excel("./tables/Section-36/yardwaste-adjustment-factors.xlsx")
df[is.na(df)] <- ""
```

```{r yardwaste-adjfactors, tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Adjustment factors for percentage of forested acres.',
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("hold_position", "scale_down"))
```

### Allocation Procedure

National values for the amount of waste generated are distributed to the counties based on rural population, as described in the previous section.

### Emission Factors

The emissions factors for CAPs are from AP-42 [[ref 7](#yardwaste-references)], the emissions inventory improvement program [[ref 8](#yardwaste-references)], and an ERTAC workgroup [[ref 9](#yardwaste-references)]. For burning of leaves, emissions factors for PM_{2.5} are calculated by multiplying the PM_{10} emissions factor by a ratio of 0.7709. Emission factors for NH3 were derived from the 2002 NEI crop residue emission estimates using the ratio of NH3/NOx for pasture grass from Pouliot et al. (2017) [[ref 10](#yardwaste-references)] and the NOx emission factor from AP-42. Emissions factors for HAPs are from an EPA Control Technology Center report [[ref 11](#yardwaste-references)].

### Controls

Controls for residential yard waste burning are generally in the form of a ban on open burning of waste in a given municipality or county. However, literature suggests that burn bans are not 100% effective. It is therefore assumed that approximately 25% of the residents that may burn yard waste would burn do so even if a ban is in place. For counties that have burn bans, the assumption is applied by multiplying 0.25 by the annual waste burned. Currently no counties are assumed to have burn bans in place. 

If county *c* has a burn ban, then:
\begin{equation} 
  LW_{c} =  LW_{c} \times 0.25
  (\#eq:leaf-control)
\end{equation}

\begin{equation} 
  BW_{c} =  BW_{c} \times 0.25
  (\#eq:brush-control)
\end{equation}
 
<div style="margin-left: 2em;">
  Where: <br> \newline
  \( LW_{c} \) = Annual leaf waste burned in county *c*, in tons <br> \newline
  \( BW_{c} \) = Annual brush waste burned in county *c*, in tons <br> \newline
</div>

### Emissions

The annual amount of combustible lead and brush waste burned in each county is multiplied by the appropriate emissions factors:

\begin{equation} 
  E_{p,c} =  EF_{p} \times LW_{c} \times \frac{1 ton}{2000 lbs}
  (\#eq:leaf-emissions)
\end{equation}

\begin{equation} 
  E_{p,c} =  EF_{p} \times BW_{c} \times \frac{1 ton}{2000 lbs}
  (\#eq:brush-emissions)
\end{equation}
 
<div style="margin-left: 2em;">
  Where: <br> \newline
  \( E_{p,c} \) = Annual emissions of pollutant *p* in county *c* in tons <br> \newline
  \( EF_{p} \) = Emissions factor for pollutant *p*, in lbs./ton of waste burned <br> \newline
  \( LW_{c} \) = Annual leaf waste burned in county *c*, in tons <br> \newline
  \( BW_{c} \) = Annual brush waste burned in county *c*, in tons <br> \newline
</div>

### Sample Calculations

Table \@ref(tab:yardwaste-sample-calculations) lists sample calculations to determine the CO emissions from open burning of yard waste. The values in these equations are demonstrating program logic and are not representative of any specific NEI year or county. 

```{r, include=FALSE}
df <- data.frame(
  Eq = c(1:8),
  Equation = c("$PC_{yw} = \\frac{YW \\times 0.60}{P_{US}}$",
               "$PBurn_{c} = RPop_{c} \\times 0.24$",
               "$LW_{c} = PBurn_{c} \\times PC_{yw} \\times YWFr_{t} \\times AF_{fa,c}$",
               "$BW_{c} = PBurn_{c} \\times PC_{yw} \\times YWFr_{t} \\times AF_{fa,c}$",
               "$LW_{c} =  LW_{c} \\times 0.25$",
               "$BW_{c} =  BW_{c} \\times 0.25$",
               "$E_{p,c} =  EF_{p} \\times LW_{c}$",
               "$E_{p,c} =  EF_{p} \\times BW_{c}$"),
  Values = c("$\\frac{34.5\\text{ million tons} \\times 0.60}{318.85\\text{ million people}}$",
             "$22,921\\text{ people} \\times 0.24$",
             "$5,501\\text{ people} \\times 0.065\\text{ tons} \\times 0.25 \\times 1$",
             "$5,501\\text{ people} \\times 0.065\\text{ tons} \\times 0.25 \\times 1$",
             "$\\text{N/A}$",
             "$\\text{N/A}$",
             "$89.39\\text{ tons leaf waste} \\times 112\\text{ lbs. per ton} \\times 0.005 \\frac{ton}{lbs.}$",  
             "$89.39\\text{ tons leaf waste} \\times 140\\text{ lbs. per ton} \\times 0.005 \\frac{ton}{lbs.}$"),  
  Result = c("0.065 tons yard waste per person per year",
             "5,501 people likely to burn waste",
             "89.39 tons of leaf waste burned",
             "89.39 tons of brush waste burned",
             "County does not have a burn ban",
             "County does not have a burn ban",
             "5.01 tons CO emissions from burning of leaf waste",
             "6.26 tons CO emissions from burning of brush waste")
)
```

```{r yardwaste-sample-calculations, tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::landscape(
  kableExtra::kbl(
    df,booktabs = TRUE,longtable = TRUE,
    escape = FALSE,  # allow LaTeX in cells
    caption = "Sample Calculations",
    col.names = c("Eq. \\#", "Equation", "Values", "Result"),
  ) %>%
    kableExtra::kable_styling(position = "center",
      latex_options = c("repeat_header"),font_size = 12) %>%
    kableExtra::column_spec(1, latex_column_spec = "p{2cm}") %>%
    kableExtra::column_spec(2, latex_column_spec = "p{7cm}") %>%
    kableExtra::column_spec(3, latex_column_spec = "p{7cm}") %>%
    kableExtra::column_spec(4, latex_column_spec = "p{5cm}")
)
```

### Improvements/Changes in the 2023 NEI

There have been no significant changes to this category.

### Puerto Rico and U.S. Virgin Islands

Emissions from Puerto Rico are calculated using the same method described above. For the U.S. Virgin Islands, emissions are calculated using 2020 population data [[ref 3](#yardwaste-references)], since 2023 Census Data does not exist for the U.S. Virgin Islands. 

## References {#yardwaste-references}

1.	U.S. EPA. 2020. [Advancing Sustainable Materials Management: 2018 Fact Sheet, Table1. Generation, Recovery, and Discards of Products in MSW, 2018.](https://www.epa.gov/sites/default/files/2021-01/documents/2018_ff_fact_sheet_dec_2020_fnl_508.pdf)
2.	U.S. Environmental Protection Agency. 2011. [Municipal Solid Waste Generation, Recycling, and Disposal in the United States: Facts and Figures for 2010—Fact Sheet](https://archive.epa.gov/epawaste/nonhaz/municipal/web/pdf/msw_2010_factsheet.pdf), p. 4.  
3.	U.S. Census Bureau. [County Population Totals.](https://www.census.gov/data/tables/time-series/demo/popest/2020s-counties-total.html)
4.  Environment Canada. 2001. “[Household Garbage Disposal and Burning](https://gaftp.epa.gov/air/nei/2023/doc/supporting_data/nonpoint/Env%20Canada%20-%20Household%20garbage%20disposal%20and%20burning.pdf).” Prepared by Environics Research Group.
5.	Two Rivers Regional Council of Public Officials and Patrick Engineering, Inc. 1994.  “Emission Characteristics of Burn Barrels,” prepared for the U.S. Environmental Protection Agency, Region V. 
6.	U.S. Environmental Protection Agency. 2023. [Biogenic Emission Inventory System (BEIS).](https://www.epa.gov/air-emissions-modeling/biogenic-emission-inventory-system-beis)
7.	U.S. Environmental Protection Agency. 1992. [AP-42, Fifth Edition, Volume 1, Chapter 2: Solid Waste Disposal. Section 2.5: Open Burning.](https://www.epa.gov/sites/default/files/2020-10/documents/c02s05.pdf)
8.	Eastern Research Group, Inc. 2001. [Emissions Inventory Improvement Program. Volume III: Chapter 16, Open Burning.](https://www.epa.gov/sites/default/files/2015-08/documents/iii16_apr2001.pdf)
9.	Huntley, Roy, 2009. Spreadsheet “state_comparison ERTAC SS_version7_4 Nov 23 2009.xls”.
10.	G. Pouliot, V. Rao, J.L. McCarty, and A. Soja. 2017. [Development of the crop residue and rangeland burning in the 2014 National Emissions Inventory using information from multiple sources.](https://pmc.ncbi.nlm.nih.gov/articles/PMC6088810/) Journal of the Air & Waste Management Association, 67(5), 613-622. 
11.	U.S. Environmental Protection Agency, Control Technology Center. 1997. [Evaluation of Emissions from the Open Burning of Household Waste in Barrels.](https://cfpub.epa.gov/si/si_public_record_Report.cfm?Lab=NRMRL&dirEntryID=115129) EPA 600/R 97 134a.