```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
library(glue)
```

# Publicly-Owned Treatment Works {#potws}

## Sector Descriptions and Overview 

This source category, Publicly Owned Treatment Works (POTW), includes treatment works owned by a state, municipality, city, town, special sewer district, or other publicly owned and financed entity as opposed to a privately (industrial) owned treatment facility. The definition includes intercepting sewers, outfall sewers, sewage collection systems, pumping, power, and other equipment. The wastewater treated by these POTWs is generated by industrial, commercial, and domestic sources [[ref 1](#potws-references)]. The SCC that EPA uses for estimated nonpoint emissions is 2630020000; the SCC description is “Waste Disposal, Treatment, and Recovery; Wastewater Treatment; Public Owned; Total Processed”.

## EPA-developed estimates

The calculations for estimating the emissions from POTWs involve multiplying the wastewater flow rate by emissions factors for VOCs, NH3, and 53 HAPs.

### Activity Data

The activity data for this source category is the wastewater flow rate. The EPA Clean Watersheds Needs Survey provides flow rate by facility and estimates the national POTW flow rate in 2012 for all facilities as 32,822 million gallons per day (MMGD) [[ref 2](#potws-references)]. The nationwide flow rate includes Puerto Rico and the US Virgin Islands. To estimate flow rates in the current inventory year, facility-level daily flow rates in 2012 are multiplied by the ratio of current year population to 2012 population in the county where the facility resides [[ref 3](#potws-references)]. County-level annual wastewater flow rates are calculated by summing the daily flow rates for all POTWs within the county and multiplying by 365 days in a year. 

\begin{equation} 
  FR_{c,y} =  \sum^{n}_{f=1}{FR_{f,2012}} \times 365 \text{days per year} \times \frac{P_{c,y}}{P_{c,2012}} 
  (\#eq:flow-rate)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( FR_{c,y} \) = The annual wastewater flow rate of county *c* in inventory year *y* <br> \newline
  \( FR_{f,2012} \) = The daily wastewater flow rate at facility *f* in 2012 <br> \newline
  \( P_{c,y} \) = Total population in county *c* in inventory year *y* <br> \newline
  \( P_{c,2012} \) =Total population in county *c* in 2012  <br> \newline
  
</div>

### Allocation Procedure

For a given county, county-level wastewater flow rates are calculated by summing the flow rates for all POTWs within the county.

### Emission Factors

The ammonia emissions factor was obtained from an EPA report [[ref 4](#potws-references)] and the VOC emissions factor was based on a Tri-County Technical Advsory Committee (Tri-TAC) study among 3 California counties [[ref 5](#potws-references)]. Emissions factors for HAPs were derived using 1996 area source emissions estimates that were provided by Bob Lucas [[ref 6](#potws-references)] and the 1996 nationwide flow rate [[ref 7](#potws-references)]. These HAP emissions factors were then multiplied by the 2008 to 2002 VOC emissions factor ratio (0.85/9.9) to obtain the final HAP emissions factors applied.

### Controls

There are no controls assumed for this category. 

### Emissions

Emissions are estimated by multiplying an emissions factor by the county flow rate. A conversion factor was used to convert pounds to tons. 

\begin{equation} 
  E_{p,c} =  FR_{c,y} \times EF_{p} \times \frac{\text{1 ton}}{\text{2,000 lbs}} 
  (\#eq:emissions-potws)
\end{equation}
 
<div style="margin-left: 2em;">
  Where: <br> \newline
  \( E_{p,c} \) = Nonpoint emissions of pollutant *p* in county *c*, in tons <br> \newline
  \( FR_{c,y} \) =Flow rate of county *c* and inventory year *y*, in MMGY <br> \newline
  \( EF_{p} \) = Emissions factor for pollutant *p*, in lbs. per MMGAL <br> \newline
</div>

### Point Source Subtraction

The county-level flow rates include all facilities reported as POTWs in the EPA Clean Watersheds Needs Survey. In some cases, SLT agencies might include facilities under their point source inventory reporting. In these cases, SLT agencies have two options for submitting state-level point source data to EPA for point source subtraction:
<br> \newline
- Option A: County-level flow rates associated with POTWs reported as point sources; or <br> \newline
-	Option B: County-level emissions of VOC and NH3 for POTWs reported as point sources. 

### Sample Calculations

Table \@ref(tab:potw-sample-calculations) lists sample calculations to determine the benzene emissions for nonpoint source POTWs. The values in these equations are demonstrating program logic and are not representative of any specific NEI year or county.

```{r, include=FALSE}
df <- data.frame(
  Eq = c(1, 2),
  Equation = c("$FR_{c,y} = \\sum^{n}_{f=1}{FR_{f,2012}} \\times 365 \\times \\frac{P_{c,y}}{P_{c,2012}}$",
               "$E_{p,c} =  FR_{c,y} \\times EF_{p} \\times \\frac{1\\text{ ton}}{2,000\\text{ lbs}}$"),
  Values = c("$2.866\\text{ MMGD} \\times 365\\text{ days} \\times \\frac{55,504\\text{ people}}{54,927\\text{ people}}$",
             "$1,057.07\\text{ MMGY} \\times 0.00673\\text{ lb per MMG} \\times \\frac{1\\text{ ton}}{2,000\\text{ lbs}}$"), 
  Result = c("1,057.07 MMGY",
             "0.003557 tons benzene per year")
)
```
```{r potw-sample-calculations, tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::landscape(
  kableExtra::kbl(
    df,booktabs = TRUE,longtable = TRUE,
    escape = FALSE,  # allow LaTeX in cells
    caption = "Sample Calculations",
    col.names = c("Eq. \\#", "Equation", "Values", "Result"),
  ) %>%
    kableExtra::kable_styling(position = "center",
      latex_options = c("repeat_header"),font_size = 10) %>%
    kableExtra::column_spec(1, latex_column_spec = "p{2cm}") %>%
    kableExtra::column_spec(2, latex_column_spec = "p{7cm}") %>%
    kableExtra::column_spec(3, latex_column_spec = "p{6cm}") %>%
    kableExtra::column_spec(4, latex_column_spec = "p{6cm}")
)
```

### Improvements/Changes in the 2023 NEI

There have been no significant changes to this category since the 2020 NEI.

### Puerto Rico and U.S. Virgin Islands

Emissions from Puerto Rico are calculated using the same method described above. For the U.S. Virgin Islands, emissions are not multiplied by the ratio of 2020 to 2012 population since 2020 Census Data does not exist for the U.S. Virgin Islands.

## References {#potws-references}

1.	U.S. EPA, 64FR57572, [National Emission Standards for Publicly Owned Treatment Works](https://www.govinfo.gov/content/pkg/FR-1999-10-26/pdf/99-27799.pdf), Final Rule, 40 CFR Part 63, 26 October 1999. 
2.	U.S. Environmental Protection Agency, [Clean Watersheds Needs Survey](https://ordspub.epa.gov/ords/cwns2012/f?p=CWNS2012:1:0:::::)  2012 Data and Reports
3.	U.S. Census Bureau. [Total Population](https://data.census.gov/table?tid=ACSDT5Y2020.B01003), American Community Survey
4.	Stephen M. Roe, Melissa D. Spivey, Holly C. Lindquist, Kirstin B. Thesing, and Randy P. Strait, E.H. Pechan & Associates, Inc., [Estimating Ammonia Emissions from Anthropogenic Nonagricultural Sources – Draft Final Report](https://www.epa.gov/sites/production/files/2015-08/documents/eiip_areasourcesnh3.pdf), prepared for U.S. Environmental Protection Agency, Emission Inventory Improvement Program, April 2004.   
5.	Prakasam Tata, Jay Witherspoon, Cecil Lue-Hing (eds.), VOC Emissions from Wastewater Treatment Plants: Characterization, Control, and Compliance, Lewis Publishers, 2003, p. 261. https://doi.org/10.1201/9780203011522
6.	Memorandum from Bob Lucas, U.S Environmental Protection Agency to Greg Nizich, U.S. Environmental Protection Agency, “Review of Baseline Emissions Inventory,” 16 October 1998.
7.	U.S. Environmental Protection Agency, [Facilities Database (Needs Survey) - Frequently Asked Questions,](https://permanent.access.gpo.gov/websites/epagov/www.epa.gov/OWM/mtb/cwns/1996rtc/faqwfd.htm) accessed 30 April 2019.